EDDA Validation by Pyrosequencing
========================================================
##### By: Sumaiya Islam
##### Date: August 6, 2016

### Script contents:
- Confirmation of differential CpG sites identified by 450K array analysis (cg07280807,cg18752527) in an independent validation. Note that these pyrosequencing assays were performed by Chris in Tubingen using a separate batch of bisulfite-converted DNA from isolated T-cell samples of these individuals. 

### A. Set up wd & packages

We will initially set our working directory and load our libraries.
```{r setup, include = FALSE, echo=FALSE}
setwd("/home/sislam/EDDA")
library(methylumi)
library(gplots)
library(marray)
library(lumi)
library(lattice)
library("RColorBrewer")
library(knitr)
library(xtable)
library(limma)
library(reshape)
library(ggplot2)
library(grid)
library(gridExtra)
library(MASS)
```

### B. Load data

We will load the pyrosequencing data for the target CpGs in validation cohort (cg07280807,cg18752527). 

```{r load_data, echo=FALSE, include=FALSE}
# load pyrosequencing data from Chris
validation.dat<-read.csv(file="EDDA_Validation_Cohort_Pyro.csv", header = T, check.names=F)
validation.dat$Group<-as.factor(validation.dat$Group)
validate.pyro.dat<-validation.dat[,c("Sample","cg07280807", "cg18752527")]
validate.pyro.melted<-melt(validate.pyro.dat, id.vars="Sample")
colnames(validate.pyro.melted)<-c("Sample", "Probe", "Pyro_value")
# merge with meta data
validation.dat.plot<-merge(validate.pyro.melted, validation.dat[,c("Sample", "Group")], by.y="Sample", by.x="Sample")

head(validation.dat.plot)
```

### B. Test if these CpG sites are significantly different using pyro measures in validation samples


Test if averaged methylation values for each assay are significantly different between groups

```{r echo=FALSE, fig.height=3, fig.width=8, warning=FALSE}

# subset by probe and select only controls and T1 patients
cg07280807.test.dat<-subset(validation.dat.plot, Probe=="cg07280807")
cg18752527.test.dat<-subset(validation.dat.plot, Probe=="cg18752527")

## test differences between controls and T1 patients one-sided T-test

(assay.1.stat<-t.test(cg07280807.test.dat[which(cg07280807.test.dat$Group=="0"),"Pyro_value"], cg07280807.test.dat[which(cg07280807.test.dat$Group=="1"),"Pyro_value"], alternative="less")) # for cg07280807 p = 0.001822

(assay.2.stat<-t.test(cg18752527.test.dat[which(cg18752527.test.dat$Group=="0"),"Pyro_value"], cg18752527.test.dat[which(cg18752527.test.dat$Group=="1"),"Pyro_value"], alternative="greater"))# for cg18752527 p = 4.51e-7


(adjusted.p.values<-p.adjust(c(assay.1.stat$p.value,assay.2.stat$p.value), method="BH"))

# place specific labels for group
groupings<-gsub("0", "Control",  gsub("1", "Patient(T1)", validation.dat.plot$Group))
validation.dat.plot$Group<-groupings

## generate boxplots for each assay

ggplot(validation.dat.plot, aes(Group, Pyro_value))+
  geom_boxplot(aes(fill=Group, outlier.shape=NA))+
  geom_point(aes(fill=Group), shape=21, size=2, position = position_jitter(w = 0.2))+
  theme_bw()+xlab("Group")+ylab("Pyrosequencing DNA methylation")+
  scale_fill_manual(values=c("#d6604d", "#4393c3"))+
  facet_wrap(~Probe)

```

We see that these CpG sites are significantly different between controls and patients at T1 based on pyrosequencing measures. Specifically, based on one-sided t-test (after multiple test correction)

* for cg07280807: Benjamini-Hochberg corrected p = 0.048
* for cg18752527: Benjamini-Hochberg corrected p = 0.048
