EDDA Blood Cell Deconvolution
========================================================
  ## By: Sumaiya Islam
  ## Date Updated: February 15, 2016
  
  ### Script contents:
  - use blood deconvolution algorithm (from minfi package) to predict blood cell type proportions 
  - check predicted cell type proportions (since this project has purified CD3+ T-cells, expect null proportions for all other blood cell types except T-lymphocytes)
  - regress out blood cell type heterogeneity from original methylation profiles
  - check corrected methylation profiles by PCA
 
   
### A. Set up working directory & packages

We will initially set our working directory and load our libraries.

```{r,include=FALSE}
library(methylumi)
library(limma)
library(sva)
library(minfi)
library(FlowSorted.Blood.450k)
library(IlluminaHumanMethylation450kmanifest)
library(quadprog)
```

### B. Load files

#### According to Chris Brueckmann (collaborator), for the DNA isolation, "we used CD3 Dynabeads (https://www.thermofisher.com/order/catalog/product/11151D) according to standard protocol for positive isolation: collected PBMCs are washed once, then transferred on 100µl prewashed Dynabeads. We keep this mixture rotating for 30min at 4°C. Afterwards, the beads are washed 3 times in Buffer (PBS pH 7,4, 0,1% BSA, 1mM EDTA). After this step, every CD3 neg. cell should be depleted. We lyse the cells directly from the beads and continue with the standard DNA extraction protocol.""

```{r}
load("EDDA.combat.RData") # Combat-corrected data
# check for NAs in data
ind<-is.row.na(betas(EDDA.combat)) # The function returns a vector of logical variables, one for each row of the matrix. The variable is TRUE if the row does not contain any missing values and FAlSE otherwise.
length(na.count<-which(ind=="FALSE")) # 0 (there are no rows that contain a NA in the raw data)
```

### C. Blood Deconvolution to predict blood cell type proportions

#### We will be using the blood deconvolution method from Koestler et al. 2013.  Blood based patterns of DNA methylation predict the underlying distribution of cell types: a validation analysis.  Epigenetics, 8(8), 816-26.  PMID: 23903776.

```{r}
# run source script for blood deconvolution (adapted from minifi package and passed down from Meg)
source("ECC2.R")
RGset = as.data.frame(betas(EDDA.combat))
cellprop <- ECC2(RGset) # this takes a while so best to run overnight

# check results of deconvolution
head(cellprop$counts)
diff<-as.data.frame(cellprop$counts)
# save(diff, file = "EDDA_Deconvoluted_Predicted_Cell_Proportions.RData")

load("EDDA_Deconvoluted_Predicted_Cell_Proportions.RData")
# integrate predicted cell proportions with existing meta data 
identical(rownames(diff), rownames(pData(EDDA.combat))) # TRUE
meta.cell<-cbind(pData(EDDA.combat)$Sample_Group, diff)
meta.cell$Sample<-rownames(meta.cell)
str(meta.cell)
head(melted.diff<-melt(meta.cell))
# plot predicted proportions of each cell type, coloured by sample_group
colnames(melted.diff)<-c("Group", "Sample", "Cell.Type", "Predicted.Proportion")

melted.diff$Group<-as.factor(melted.diff$Group)


ggplot(melted.diff, aes(Group, Predicted.Proportion)) + geom_boxplot(aes(fill=Group),color="black", outlier.shape=NA)+ geom_point(shape=21, fill="grey",colour="black", size=2, position=position_jitter(w=0.2)) + scale_fill_manual(values=c("#d6604d","#4393c3", "#92c5de")) + theme_bw() +facet_wrap(~Cell.Type) + xlab("Sample Group") + ylab("Predicted Cell Proportions")
```

We can see that there is notable contamination of the T-lymphocyte samples with NK cells. We will test if the cell type proportions associate with sample group. 

#### Comparison of cell type proportions between controls and T1 patient samples

```{r, echo=FALSE}
# subset controls and patient T1 samples
all.dat<-EDDA.combat
pData(all.dat)$Sample_Group<-as.character(pData(all.dat)$Sample_Group)
pData(all.dat)$Sample_Group<-gsub("patient_t2", "t2", all.dat$Sample_Group)
toMatch<-c("control", "patient")
con.t1.samples <- grep(paste(toMatch,collapse="|"),all.dat$Sample_Group)
Con_T1_subset<-all.dat[,con.t1.samples]
meta.CT1<-pData(Con_T1_subset)
diff.CT1<-diff[which(rownames(diff)%in%rownames(meta.CT1)),]
identical(rownames(diff.CT1), rownames(meta.CT1)) #TRUE 
meta.CT1.diff<-cbind(meta.CT1$Sample_Group, diff.CT1)
colnames(meta.CT1.diff)<-c("Group", "CD8T", "CD4T", "NK", "Bcell", "Mono", "Gran")
head(melted.diff.CT1<-melt(meta.CT1.diff))
colnames(melted.diff.CT1)<-c("Group", "Cell.Type", "Predicted.Proportion")
melted.diff.CT1$Group<-as.factor(melted.diff.CT1$Group)
ggplot(melted.diff.CT1, aes(Group, Predicted.Proportion)) + geom_point(aes(colour=Group), size=3,shape=19, position=position_jitter(w=0.2)) + scale_colour_manual(values=c("#ef8a62","#998ec3", "#5ab4ac")) + theme_bw() +facet_wrap(~Cell.Type)
# test difference between groups to see if there is a significant difference in cell type proportion between controls and T1 patient (Mann-Whitney U test)
mw.test.CD8T<-wilcox.test(meta.CT1.diff[which(meta.CT1.diff$Group=="control"), "CD8T"],meta.CT1.diff[which(meta.CT1.diff$Group=="patient"), "CD8T"])
mw.test.CD4T<-wilcox.test(meta.CT1.diff[which(meta.CT1.diff$Group=="control"), "CD4T"],meta.CT1.diff[which(meta.CT1.diff$Group=="patient"), "CD4T"])
mw.test.NK<-wilcox.test(meta.CT1.diff[which(meta.CT1.diff$Group=="control"), "NK"],meta.CT1.diff[which(meta.CT1.diff$Group=="patient"), "NK"])
mw.test.Bcell<-wilcox.test(meta.CT1.diff[which(meta.CT1.diff$Group=="control"), "Bcell"],meta.CT1.diff[which(meta.CT1.diff$Group=="patient"), "Bcell"])
mw.test.Mono<-wilcox.test(meta.CT1.diff[which(meta.CT1.diff$Group=="control"), "Mono"],meta.CT1.diff[which(meta.CT1.diff$Group=="patient"), "Mono"])
mw.test.Gran<-wilcox.test(meta.CT1.diff[which(meta.CT1.diff$Group=="control"), "Gran"],meta.CT1.diff[which(meta.CT1.diff$Group=="patient"), "Gran"])
# multiple test correction
p.vals<-c(mw.test.CD8T$p.value, mw.test.CD4T$p.value, mw.test.NK$p.value, mw.test.Bcell$p.value, mw.test.Mono$p.value, mw.test.Gran$p.value)
CT1.test.output<-as.matrix(adjusted.pvals.CT1<-p.adjust(p.vals, method="BH"))
rownames(CT1.test.output)<-c("CD8T", "CD4T", "NK", "Bcell", "Mono", "Gran")
colnames(CT1.test.output)<-"adjusted.pval"
CT1.test.output
```

There seems to be a statistically different proportions of granulocytes between control and T1 patient samples (adjusted p-value < 0.05)---however, that may be due to the low overall variability in granulocyte proportions (note that a warning was provided saying cannot compute p-values with ties).

#### Comparison of cell type proportions between T1 and T2 patient samples
```{r, echo=FALSE}
# subset data to just T1 vs T2 patient samples
toMatch2<-c("patient", "t2")
tt.samples <- grep(paste(toMatch2,collapse="|"),all.dat$Sample_Group)
tt_subset<-all.dat[,tt.samples]
meta.tt<-pData(tt_subset)
meta.tt$Sample_Group<-as.factor(as.character(meta.tt$Sample_Group))
meta.tt$Sample_Name<-colnames(exprs(tt_subset))
meta.tt$Patient_ID<-c("69", "66", "103", "67", "89", "84", "120", "47", "85", "47", "49", "123", "31", "31", "107", "122", "87", "67", "123", "87", "103", "69", "61", "83", "66", "49", "104", "121", "61", "121", "107", "64","83", "104", "100", "105", "82", "82", "105", "68", "100", "89", "84", "68", "120", "85", "64", "122")
meta.tt$Patient_ID<-as.factor(meta.tt$Patient_ID)

# combine meta data with predicted cell proportions
diff.tt<-diff[which(rownames(diff)%in%rownames(meta.tt)),]
identical(rownames(diff.tt), rownames(meta.tt)) #TRUE 
meta.tt.diff<-cbind(meta.tt$Sample_Group,meta.tt$Patient_ID, diff.tt)
colnames(meta.tt.diff)<-c("Group", "Patient.ID","CD8T", "CD4T", "NK", "Bcell", "Mono", "Gran")
head(melted.diff.tt<-melt(meta.tt.diff))
colnames(melted.diff.tt)<-c("Group", "Patient.ID", "Cell.Type", "Predicted.Proportion")
melted.diff.tt$Group<-as.factor(melted.diff.tt$Group)
# plot predicted proportions for each cell type by group
ggplot(melted.diff.tt, aes(Group, Predicted.Proportion)) + geom_point(aes(colour=Group), size=3,shape=19, position=position_jitter(w=0.2)) + scale_colour_manual(values=c("#66c2a5","#fc8d62")) + theme_bw() +facet_wrap(~Cell.Type)
# test difference between groups to see if there is a significant difference in cell type proportion between T1 and T2 patient (Wilcoxon signed rank test)
meta.T1.diff<-meta.tt.diff[which(meta.tt.diff$Group=="patient"),]
meta.T2.diff<-meta.tt.diff[which(meta.tt.diff$Group=="t2"),]
meta.T2.diff.order<-meta.T2.diff[match(meta.T1.diff$Patient.ID, meta.T2.diff$Patient.ID),]
identical(meta.T1.diff$Patient.ID, meta.T2.diff.order$Patient.ID)
mw.test.CD8T<-wilcox.test(meta.T1.diff$CD8T, meta.T2.diff.order$CD8T, paired=TRUE)
mw.test.CD4T<-wilcox.test(meta.T1.diff$CD4T, meta.T2.diff.order$CD4T, paired=TRUE)
mw.test.NK<-wilcox.test(meta.T1.diff$NK, meta.T2.diff.order$NK, paired=TRUE)
mw.test.Bcell<-wilcox.test(meta.T1.diff$Bcell, meta.T2.diff.order$Bcell, paired=TRUE)
mw.test.Mono<-wilcox.test(meta.T1.diff$Mono, meta.T2.diff.order$Mono, paired=TRUE)
mw.test.Gran<-wilcox.test(meta.T1.diff$Gran, meta.T2.diff.order$Gran, paired=TRUE)
p.vals<-c(mw.test.CD8T$p.value, mw.test.CD4T$p.value, mw.test.NK$p.value, mw.test.Bcell$p.value, mw.test.Mono$p.value, mw.test.Gran$p.value)
# multiple test correction
tt.test.output<-as.matrix(adjusted.pvals.tt<-p.adjust(p.vals, method="BH"))
rownames(tt.test.output)<-c("CD8T", "CD4T", "NK", "Bcell", "Mono", "Gran")
colnames(tt.test.output)<-"adjusted.pval"
tt.test.output
```

There does not seem to be any association between predicted proportion of each cell type and group status (T1 vs T2 in patient samples).

#### Comparison of cell type proportions between controls and T2 patient samples
```{r, echo=FALSE}
# subset data to just control vs T2 patient samples
toMatch3<-c("control", "t2")
c2.samples <- grep(paste(toMatch3,collapse="|"),all.dat$Sample_Group)
c2_subset<-all.dat[,c2.samples]
meta.c2<-pData(c2_subset)
meta.c2$Sample_Group<-as.factor(as.character(meta.c2$Sample_Group))
meta.c2$Sample_Name<-colnames(exprs(c2_subset))

# combine meta data with predicted cell proportions
diff.c2<-diff[which(rownames(diff)%in%rownames(meta.c2)),]
identical(rownames(diff.c2), rownames(meta.c2)) #TRUE 
meta.c2.diff<-cbind(meta.c2$Sample_Group, diff.c2)
colnames(meta.c2.diff)<-c("Group","CD8T", "CD4T", "NK", "Bcell", "Mono", "Gran")
head(melted.diff.c2<-melt(meta.c2.diff))
colnames(melted.diff.c2)<-c("Group", "Cell.Type", "Predicted.Proportion")
melted.diff.c2$Group<-as.factor(melted.diff.c2$Group)
# plot predicted proportions for each cell type by group
ggplot(melted.diff.c2, aes(Group, Predicted.Proportion)) + geom_point(aes(colour=Group), size=3,shape=19, position=position_jitter(w=0.2)) + scale_colour_manual(values=c("#d53e4f","#3288bd")) + theme_bw() +facet_wrap(~Cell.Type)

# test difference between groups to see if there is a significant difference in cell type proportion between controls and T2 patient (Mann-Whitney U test)
mw.test.CD8T<-wilcox.test(meta.c2.diff[which(meta.c2.diff$Group=="control"), "CD8T"],meta.c2.diff[which(meta.c2.diff$Group=="t2"), "CD8T"])
mw.test.CD4T<-wilcox.test(meta.c2.diff[which(meta.c2.diff$Group=="control"), "CD4T"],meta.c2.diff[which(meta.c2.diff$Group=="t2"), "CD4T"])
mw.test.NK<-wilcox.test(meta.c2.diff[which(meta.c2.diff$Group=="control"), "NK"],meta.c2.diff[which(meta.c2.diff$Group=="t2"), "NK"])
mw.test.Bcell<-wilcox.test(meta.c2.diff[which(meta.c2.diff$Group=="control"), "Bcell"],meta.c2.diff[which(meta.c2.diff$Group=="t2"), "Bcell"])
mw.test.Mono<-wilcox.test(meta.c2.diff[which(meta.c2.diff$Group=="control"), "Mono"],meta.c2.diff[which(meta.c2.diff$Group=="t2"), "Mono"])
mw.test.Gran<-wilcox.test(meta.c2.diff[which(meta.c2.diff$Group=="control"), "Gran"],meta.c2.diff[which(meta.c2.diff$Group=="t2"), "Gran"])

# multiple test correction
p.vals<-c(mw.test.CD8T$p.value, mw.test.CD4T$p.value, mw.test.NK$p.value, mw.test.Bcell$p.value, mw.test.Mono$p.value, mw.test.Gran$p.value)
c2.test.output<-as.matrix(adjusted.pvals.c2<-p.adjust(p.vals, method="BH"))
rownames(c2.test.output)<-c("CD8T", "CD4T", "NK", "Bcell", "Mono", "Gran")
colnames(c2.test.output)<-"adjusted.pval"
c2.test.output
```

There does not seem to be any association between predicted proportion of each cell type and group status (controls vs T2 in patient samples).

### D. Regress out inter-individual differences in blood cell type proportions from DNA methylation data

```{r echo=FALSE}
identical(colnames(betas(all.dat)), rownames(diff)) # Should be TRUE
betas<-betas(all.dat) 
# fit linear model for cell type effects on each probe in betas
betas.lm<-apply(betas, 1, function(x){
  blood<-diff[colnames(betas),]
  lm(x~CD8T+CD4T+NK+Bcell+Mono+Gran,data=blood)
})
# extract matrix of residuals from resulting linear models
residuals<-t(sapply(betas.lm, function(x)residuals(summary(x))))
colnames(residuals)<-colnames(betas) # re-name residuals columns with sample names
# add the residuals of each regression model to the mean methylation value of each probe (mean across all samples) to obtain the âadjustedâ methylation data.
adj.residuals<-residuals+matrix(apply(betas, 1, mean), nrow=nrow(residuals), ncol=ncol(residuals))
head(adj.residuals)

# To make sure we do not induce any NAs into the dataset when we convert the beta values back M-values (by log2 transformation), we need to ensure we do not have any corrected beta values that are greater or equal to zero or any beta values that are greater than 1. 

adj.residuals[adj.residuals<=0]<-0.001 # convert any values that are less than or equal to zero to 0.001
adj.residuals[adj.residuals>1]<-0.999 # convert any values that are greater than 1 to 0.999
adj.M.values<-beta2m(adj.residuals)
any(is.na(adj.M.values)) # should be FALSE indicating there are no NAs
summary(adj.residuals)

# replace adjusted data into a cloned methylumi object
EDDA.cell.cor<-all.dat
identical(rownames(adj.M.values), rownames(exprs(EDDA.cell.cor))) # should be TRUE for probe order check
identical(colnames(adj.M.values), colnames(exprs(EDDA.cell.cor))) # should be TRUE for sample order check
exprs(EDDA.cell.cor)<-adj.M.values
all.equal(exprs(EDDA.cell.cor),adj.M.values) # TRUE
# save(EDDA.cell.cor,file = "EDDA.cell.cor.RData")
load("EDDA.cell.cor.RData")
```

### D. PCA to check DNA methylation Profiles before and after cell type correction

#### Define heat scree PCA function

```{r, echo = FALSE}

### Function of association meta variable with PC (ANOVA)
heat_scree_plot<-function(Loadings, Importance, Num, Order){
  adjust<-1-Importance[1]
  pca_adjusted<-Importance[2:length(Importance)]/adjust
  pca_df<-data.frame(adjusted_variance=pca_adjusted, PC=seq(1:length(pca_adjusted)))
  
  scree<-ggplot(pca_df[which(pca_df$PC<Num),],aes(PC,adjusted_variance))+geom_bar(stat = "identity",color="black",fill="grey")+theme_bw()+
        theme(axis.text = element_text(size =12),
              axis.title = element_text(size =15),
              plot.margin=unit(c(1,1.5,0.2,2.25),"cm"))+ylab("Variance")+
    scale_x_continuous(breaks = seq(1,Num,1))
  
  #### Heat
  ## correlate meta with PCS
  ## Run anova of each PC on each meta data variable

  aov_PC_meta<-lapply(1:ncol(meta_categorical), function(covar) sapply(1:ncol(Loadings), function(PC) summary(aov(Loadings[,PC]~meta_categorical[,covar]))[[1]]$"Pr(>F)"[1]))
   cor_PC_meta<-lapply(1:ncol(meta_continuous), function(covar) sapply(1:ncol(Loadings), function(PC) (cor.test(Loadings[,PC],as.numeric(meta_continuous[,covar]),alternative = "two.sided", method="spearman", na.action=na.omit)$p.value)))
  names(aov_PC_meta)<-colnames(meta_categorical)
  names(cor_PC_meta)<-colnames(meta_continuous)
  aov_PC_meta<-do.call(rbind, aov_PC_meta)
  cor_PC_meta<-do.call(rbind, cor_PC_meta)
  aov_PC_meta<-rbind(aov_PC_meta, cor_PC_meta)
  aov_PC_meta<-as.data.frame(aov_PC_meta)
  #adjust
  aov_PC_meta_adjust<-aov_PC_meta[,2:ncol(aov_PC_meta)]
  
    
  #reshape
  avo<-aov_PC_meta_adjust[,1:(Num-1)]
  avo_heat_num<-apply(avo,2, as.numeric)
  avo_heat<-as.data.frame(avo_heat_num)
  colnames(avo_heat)<-sapply(1:(Num-1), function(x) paste("PC",x, sep=""))
  avo_heat$meta<-rownames(avo)
  avo_heat_melt<-melt(avo_heat, id=c("meta"))
  
  # cluster meta data
  ord <- Order
  meta_var_order<-unique(avo_heat_melt$meta)[rev(ord)]
  avo_heat_melt$meta <- factor(avo_heat_melt$meta, levels = meta_var_order)
  
  # color if sig
  # avo_heat_melt$Pvalue<-sapply(1:nrow(avo_heat_melt), function(x) if(avo_heat_melt$value[x]>=0.9){">=0.9"}else{
   # if(avo_heat_melt$value[x]>=0.5){">=0.5"}else{
     # if(avo_heat_melt$value[x]>=0.1){">=0.1"}else{"<0.1"}}})
  avo_heat_melt$Pvalue<-sapply(1:nrow(avo_heat_melt), function(x) if(avo_heat_melt$value[x]<=0.001){"<=0.001"}else{
     if(avo_heat_melt$value[x]<=0.01){"<=0.01"}else{
       if(avo_heat_melt$value[x]<=0.05){"<=0.05"}else{">0.05"}}})
  
  heat<-ggplot(avo_heat_melt, aes(variable,meta, fill = Pvalue)) +
  geom_tile(color = "black",size=0.5) +
  theme_gray(8)+scale_fill_manual(values=c("#084594","#4292c6","#9ecae1","#deebf7"))+
      theme(axis.text = element_text(size =10, color="black"),
            axis.text.x = element_text(),
          axis.title = element_text(size =15),
          legend.text = element_text(size =14),
          legend.title = element_text(size =12),
          legend.position = c(1, 0), legend.justification = c(1,0),
          plot.margin=unit(c(0,2.25,1,1),"cm"))+
    xlab("Principal Component")+ylab(NULL)
  
  grid.arrange(scree, heat, ncol=1)
}
```

#### Pull out uncorrected PBMC methylation profiles and associated metadata
```{r echo=FALSE}

## re-structure meta data: change categorical variables to factors for ANOVA and continuous variables to numeric for Spearman's correlation
EDDA.Des<-pData(EDDA.combat)
identical(rownames(diff), rownames(EDDA.Des)) #TRUE 
EDDA.Des$nationality<-as.factor(EDDA.Des$nationality)
EDDA.Des$Sentrix_ID<-as.factor(EDDA.Des$Sentrix_ID)
EDDA.Des$Sample_Group<-as.factor(EDDA.Des$Sample_Group)
for (i in 1:nrow(EDDA.Des)){
  EDDA.Des$Row[i]<-paste(substr(EDDA.Des[i,"Sentrix_Position"], start=1, stop=3))
}
EDDA.Des$Row<- as.factor(EDDA.Des$Row)
meta.all<-cbind(EDDA.Des, diff) # add cell type proportions
str(meta.all)
```

#### Run PCA on uncorrected methylation profiles
```{r warning=FALSE, fig.height=9, fig.width=11, echo=FALSE}
betas.uncorr<-betas(EDDA.combat)
## PCA
PCA_full<-princomp(betas.uncorr[complete.cases(betas.uncorr),]) # scaling is not necessary for normalized dataset
Loadings<-as.data.frame(unclass(PCA_full$loadings))
vars <- PCA_full$sdev^2
Importance<-vars/sum(vars)
adjust<-1-Importance[1]
pca_adjusted<-Importance[2:length(Importance)]/adjust
(pca_df<-data.frame(adjusted_variance=pca_adjusted, PC=seq(1:length(pca_adjusted))))
sum(pca_df$adjusted_variance[1:11]) # first 11 PCs explain 60% of variance
# save(pca_df, file="Adj_PC_variance_cortexDat.uncor.txt")
#Specify which covariates are categorical and/or categorical
colnames(meta.all)
meta_categorical<-meta.all[,c("Sentrix_ID", "Row", "Sample_Group")]  # input column numbers in meta that contain categorical variables
meta_continuous<-meta.all[,c("CD8T","CD4T","NK","Bcell","Mono","Gran","Age", "Cigarettes_perDay")] # input column numbers in meta that contain continuous variables
meta_continuous<-data.frame(meta_continuous)
colnames(meta_categorical)<-c("Sentrix_ID", "Row", "Sample_Group")
colnames(meta_continuous)<-c("CD8T","CD4T","NK","Bcell","Mono","Gran","Age", "Cigarettes_perDay")

# Specify the number of PCs you want shown (usually # of samples in the dataset)
Num<-11

# Designate what order you want the variables to appear (continuous variables rbinded to categorical variables in function)
Order<-1:11

#Apply function on PCA results, pulls in the meta data and beta values from above
heat_scree_plot(Loadings, Importance, Num, Order)
```

The blood cell types  come up in the first PC signifying that they are attributed to a significant amount of DNA methylation variation and may be associated with age and/or cigarettes per Day. Group and CD8T and Bcell all co-occur in PC5.  

#### Run PCA on corrected PBMC methylation profiles
```{r warning=FALSE, fig.height=9, fig.width=11, echo=FALSE}
identical(rownames(meta.all), colnames(exprs(EDDA.cell.cor))) #TRUE
betas.cor<-betas(EDDA.cell.cor)
## PCA
PCA_full<-princomp(betas.cor[complete.cases(betas.cor),]) # scaling is not necessary for normalized dataset
Loadings<-as.data.frame(unclass(PCA_full$loadings))
vars <- PCA_full$sdev^2
Importance<-vars/sum(vars)
adjust<-1-Importance[1]
pca_adjusted<-Importance[2:length(Importance)]/adjust
(pca_df<-data.frame(adjusted_variance=pca_adjusted, PC=seq(1:length(pca_adjusted))))
sum(pca_df$adjusted_variance[1:11]) # first 11 PCs explain 45% of variance
# save(pca_df, file="Adj_PC_variance_cortexDat.uncor.txt")

#Specify which covariates are categorical and/or categorical
colnames(meta.all)
meta_categorical<-meta.all[,c("Sentrix_ID", "Row", "Sample_Group")]  # input column numbers in meta that contain categorical variables
meta_continuous<-meta.all[,c("CD8T","CD4T","NK","Bcell","Mono","Gran","Age", "Cigarettes_perDay")] # input column numbers in meta that contain continuous variables
meta_continuous<-data.frame(meta_continuous)
colnames(meta_categorical)<-c("Sentrix_ID", "Row", "Sample_Group")
colnames(meta_continuous)<-c("CD8T","CD4T","NK","Bcell","Mono","Gran","Age", "Cigarettes_perDay")

# Specify the number of PCs you want shown (usually # of samples in the dataset)
Num<-11

# Designate what order you want the variables to appear (continuous variables rbinded to categorical variables in function)
Order<-1:11

#Apply function on PCA results, pulls in the meta data and beta values from above
heat_scree_plot(Loadings, Importance, Num, Order)
```

We can see that Sample_Group is now associated with PC1 in the corrected methylation data, along with Age.

We will re-run differential DNA methylation analysis using the corrected data. 