HECW2 Pyrosequencing in Whole Blood
========================================================
##### By: Sumaiya Islam
##### Date: November 3, 2016

### Script contents:
- Analysis of differential CpG sites in HECW2 measured in whole blood in both the original EDDA cohort and in a separate validation cohort 

### A. Set up wd & packages

We will initially set our working directory and load our libraries.
```{r setup, include = FALSE, echo=FALSE}
setwd("/home/sislam/EDDA")
library(methylumi)
library(gplots)
library(marray)
library(lumi)
library(lattice)
library("RColorBrewer")
library(knitr)
library(xtable)
library(limma)
library(reshape)
library(ggplot2)
library(grid)
library(gridExtra)
library(MASS)
library(wateRmelon)
```

### B. Load data

We will load the pyrosequencing data for the target CpGs in validation cohort (cg07280807,cg18752527). 

```{r load_data, echo=FALSE, include=FALSE}
# load pyrosequencing data of original cohort from Chris
HECW2.replication.dat<-read.csv(file="HECW2_Whole_blood_Replication_Pyro.csv", header = T, check.names=F)

# place specific labels for group
groupings<-gsub("0", "Control",  gsub("1", "Patient(T1)", HECW2.replication.dat$Group))
HECW2.replication.dat$Group<-groupings
HECW2.replication.dat$Group<-as.factor(HECW2.replication.dat$Group)

# pull out columns of interest (Sample, Group, cg18752527)
ori.dat<-HECW2.replication.dat[,c("Sample","Group", "cg18752527")]
ori.dat$Cohort<-"Original"

# load pyrosequencing data of original cohort from Chris
HECW2.validation.dat<-read.csv(file="HECW2_Whole_blood_Validation_Pyro.csv" , header = T, check.names=F)

# place specific labels for group
CT1.groups<-c("0", "1")
HECW2.validation.dat.CT1<-HECW2.validation.dat[which(HECW2.validation.dat$Group%in%CT1.groups),] # subset to only controls and T1 patients
groupings<-gsub("0", "Control",  gsub("1", "Patient(T1)", HECW2.validation.dat.CT1$Group))
HECW2.validation.dat.CT1$Group<-groupings
HECW2.validation.dat.CT1$Group<-as.factor(HECW2.validation.dat.CT1$Group)
vali.dat<-HECW2.validation.dat.CT1
vali.dat$Cohort<-"Validation"

# check that column names match between the dataframes
identical(colnames(ori.dat), colnames(vali.dat)) # TRUE

# rbind the two dataframes
HECW2.whole.blood.pyro.dat<-rbind(ori.dat, vali.dat)
HECW2.whole.blood.pyro.dat$Cohort<-as.factor(HECW2.whole.blood.pyro.dat$Cohort)
head(HECW2.whole.blood.pyro.dat)
```

### B. Test if these CpG sites are significantly different using pyro measures in validation samples


Test if averaged methylation values for each assay are significantly different between groups

```{r echo=FALSE, fig.height=3, fig.width=8, warning=FALSE}

## test differences between controls and T1 patients using two-sided t-test

# for original cohort
(ori.dat.stat<-t.test(ori.dat[which(ori.dat$Group=="Control"),"cg18752527"], ori.dat[which(ori.dat$Group=="Patient(T1)"),"cg18752527"])) # for original cohort p = 0.02005

# for validation cohort
(vali.dat.stat<-t.test(vali.dat[which(vali.dat$Group=="Control"),"cg18752527"], vali.dat[which(vali.dat$Group=="Patient(T1)"),"cg18752527"])) # for validation cohort p = 6.626e-05



(adjusted.p.values<-p.adjust(c(ori.dat.stat$p.value,vali.dat.stat$p.value), method="BH"))



## generate boxplots for each assay
head(HECW2.whole.blood.pyro.dat)

ggplot(HECW2.whole.blood.pyro.dat, aes(Group, cg18752527))+
  geom_boxplot(aes(fill=Group, outlier.shape=NA))+
  geom_point(aes(fill=Group), shape=21, size=2, position = position_jitter(w = 0.2))+
  theme_bw()+xlab("Group")+ylab("cg18752527 DNA methylation (Pyro)")+
  scale_fill_manual(values=c("#d6604d", "#4393c3"))+
  facet_wrap(~Cohort)

```

We see that cg18752527 is significantly different between controls and patients at T1 based on pyrosequencing measures. Specifically, based on two-sided t-test (after multiple test correction)

* for original cohort: Benjamini-Hochberg corrected p = 0.020
* for validation cohort: Benjamini-Hochberg corrected p = 0.0001

##### Test if cg18752527 DNA methylation is T-cell specific 

Using the Reinius 450K data set of purified blood cell types, we will test to see if cg18752527 DNA methylation is associated with T-cells:

```{r echo=FALSE}
library(FlowSorted.Blood.450k)


# load Reinius data and background-subtract/colour-correct data
MSet.adult <- preprocessIllumina(FlowSorted.Blood.450k, bg.correct = TRUE, normalize = "controls", reference = 2)
Reinius.dat.betas<- betas(MSet.adult)
Reinius.meta<-pData(MSet.adult)
# save(Reinius.meta, file = "Reinius_Meta_Data.RData")
head(Reinius.dat.betas)
dim(Reinius.dat.betas)


# BMIQ normalization
load("EDDA.RData") # load EDDA methylumi object with all 485577 probes
filtered.EDDA<-EDDA[rownames(Reinius.dat.betas),] # filter to just probes in Reinius data
target.probe.order<-featureNames(filtered.EDDA)
Reinius.dat.betas.ordered<-Reinius.dat.betas[match(target.probe.order, rownames(Reinius.dat.betas)),] # match probe order
identical(rownames(exprs(filtered.EDDA)), rownames(Reinius.dat.betas.ordered)) # check probe order matches; TRUE
head(probe_design<-as.character(fData(filtered.EDDA)$INFINIUM_DESIGN_TYPE))
probe_design.v<- replace(probe_design, probe_design=="I", 1)
probe_design.cor<- replace(probe_design.v, probe_design.v=="II", 2)
probe_design.cor<-as.numeric(probe_design.cor)
identical(nrow(Reinius.dat.betas.ordered), length(probe_design.cor)) # TRUE
Reinius_betas_normalized<-apply(Reinius.dat.betas.ordered, 2, function(x) BMIQ(x,probe_design.cor)) # this code takes some time

# extract normalized beta values and reshape
betas_normalized_betas<-lapply(1:length(Reinius_betas_normalized), function(x) Reinius_betas_normalized[[x]]$nbeta)
betas_normalized_betas<-do.call(rbind, betas_normalized_betas)
betas_norm.fin<-t(betas_normalized_betas)
colnames(betas_norm.fin)<-colnames(Reinius.dat.betas.ordered)
head(betas_norm.fin)
Reinius_betas_normalized_fin<-betas_norm.fin
identical(rownames(Reinius.meta), colnames(Reinius_betas_normalized_fin)) # sample names in meta and beta data match; TRUE

# save(Reinius_betas_normalized_fin, file = "Reinius_betas_BMIQ_normalized.RData")
probe.of.interest<-c("cg18752527")
cg18752527.betas<-melt(Reinius_betas_normalized_fin[probe.of.interest,])
cg18752527.betas$Sample<-rownames(cg18752527.betas)
identical(Reinius.meta$Sample_Name, cg18752527.betas$Sample) # TRUE
plot.site<-cbind(cg18752527.betas, Reinius.meta)

# plot cg18752527 DNA methylation in relation to blood cell type
ggplot(plot.site, aes(CellTypeLong, value))+
  geom_boxplot(aes(fill=CellTypeLong), outlier.shape=NA)+
  geom_point(aes(fill=CellTypeLong), shape=21, size=2, position = position_jitter(w = 0.2))+
  theme_bw()+xlab("Blood Cell Type")+ylab("cg18752527 DNA methylation") +  guides(fill=guide_legend(title="Blood Cell Type")) + theme(axis.text.x = element_text(angle = 60, hjust=1))

# statistics for significance of association between DNAm and blood cell group
head(plot.site)
result<-aov(value~CellTypeLong, data = plot.site)
summary(result) # pvalue < 2 e-16

# see if this matches this probe's association with blood cell composition as listed in the package's table of associations
data("FlowSorted.Blood.450k.compTable")
(probe.result.tab<-FlowSorted.Blood.450k.compTable["cg18752527",]) # p.value  = 7.6 e -15
```

Using a previous 450K dataset of purified blood cell types (Reinius et al. 2012 PLoS One), we confirmed that the DNAm status of this CpG site is highly associated with T-lymphocytes, along with NK cells, signifying that the DNAm differences we measured in whole blood were driven, in part, by T-cells.  